### Car Market Exploratory Data Analysis.

#### objectives

**Libraries**

```{r}
library(dplyr) #Data Manipulation
library(ggplot2) # Data Visualization 
library(tidyr) #Data Tidying and Reshaping
library(scales) #Scale Balancing
library(ggcorrplot)
```

Data Collection

```{r}
all_cars <- readr::read_csv("combined_cars.csv")
```

Data Information

Data size

```{r}
dim(combined_cars)

str(combined_cars)
```

Data structure

```{r}
lapply(colnames(combined_cars), is.character)
```

**Data Insight to generate**

these are the following insight we are going to generate:

-   Distribution of Number of Car Owner.

-   Average Selling Price by Car Owner.

-   Distribution of Fuel Type.

-   Distribution of Price.

-   Relationship Of Number of Previous Owners

    #### Distribution Of Numbers Of Previous Owners

    ```{r}
    owner_dist <- combined_cars %>% group_by(no_own) %>% summarise(owner_count = n()) %>% arrange(desc(owner_count))

    #color
    colors <- colorRampPalette(c("#FFFF00", "#FF0000"))(nrow(owner_dist))

    barplot(owner_dist$owner_count, names.arg = owner_dist$no_own, main = "Distribution of Number Of Previous owners",
            xlab = " Numbers of owners ", ylab = "Count",
            col = colors)
    abline(h = mean(owner_dist$owner_count), lty = 3 , lwd =3)
    ```

Average selling price of cars owners

```{r}
owner_price <- combined_cars %>% group_by(no_own) %>% summarise(avg_price = mean(Selling_price)) %>% arrange(desc(avg_price))

#colors
colors_blyl <- colorRampPalette(c("#0000FF","#ADD8E6"))(nrow(owner_price))
options(scipen = 999)

#barplot
barplot(owner_price$avg_price, names.arg = owner_price$no_own, main = "Distribution Of Prices by Numbers of Previous Owners",
        xlab = "Number of Owners", ylab = "Price",
        col = colors_blyl, ylim = c(0, max(owner_price$avg_price)* 1.1))
abline(h = mean(owner_price$avg_price),lyt = 3, lwd = 3 )
```

Group by fuel type and count occurance

```{r}
fuel_distribution <- combined_cars%>%
  group_by(fuel_type)%>%
  summarise(fuel_count = n())%>%
  arrange(desc(fuel_count))

# Create a bar plot with colors
colors<- colorRampPalette(c("#ff5733", "#ffc300", "#DAf7A6"))(nrow(fuel_distribution))


#create a bar plot with colors
barplot(fuel_distribution$fuel_count,
        names.arg = fuel_distribution$fuel_type,
        col = colors,
        main = "Distribution of Fuel Type",
        xlab = "Fuel Type",
        ylab = "Count")

```

Average selling price by fuel type

```{r}
fuel_price <- combined_cars%>%
  group_by(fuel_type)%>%
  summarise(avg_price = round(mean(Selling_price), 0))%>%
  arrange(desc(avg_price))

#define color gradient
colors_blyl <- colorRampPalette(c("#0000FF", "#ADD8E6"))(nrow(fuel_price))


#plotting the barplot 
barplot(fuel_price$avg_price,
        names.arg = fuel_price$fuel_type,
        col = colors_blyl,
        main = "Average selling price by Fuel Type",
        xlab = "fuel type",
        ylab = "Average price",
        ylim = c(0, max(fuel_price$avg_price)*1.4)) # use avg_price for y_axis limit
 
#adding a horizontal line at the overall average price
abline(h = mean(fuel_price$avg_price),lty = 2,lwd = 2)

options(scipens = 999)

ggplot(fuel_price, aes(x = reorder(fuel_type,-avg_price), y = avg_price ))+ geom_bar(stat = "identity", fill = "lightblue", color = "black") + 
  labs(title = "Average selling price by fuel type", 
       x = "Fuel type",
       y = "Average price")+
  theme_minimal()
```

price Distribution

```{r}
#distribution of force 
ggplot(combined_cars, aes(x = Selling_price))+
  geom_histogram(bins = 45,fill = "lightblue",colors = "black")+
  geom_vline(xintercept = mean(combined_cars$Selling_price),
             linetype = "dashed", color = "red", lwd = 1)+
  geom_vline(xintercept = median(combined_cars$Selling_price),
             linetype = "dashed", color = "blue", lty = 3, lwd = 1 )+
  labs(title = "Distribution of car Price", x = "selling_price", y = "count")+
  theme_minimal()
```

Relationship Between kilometer Driven and Price

```{r}
ggplot(combined_cars, aes(x = Year_mfd, y = Selling_price, colors = fuel_type))+
  geom_point(alpha = 1.5)+
  labs(title = "Price vs Year maunfactured",
       y = "selling_price")+
  theme_bw()
```

Correlation between price and kilometer driven

```{r}
#scatter plot to show the relationship between price and kilometer driven 
ggplot(combined_cars, aes(x = km_drv,y= Selling_price))+
  geom_point(alpha = 0.5, color = "blue")+ #point with some transparency
  labs(title = "Correlation Between Price And Kilometer Driven",
       x = "Kilometer driven",
       y = "selling price")+
  theme_minimal()

```

```{r}
cor_matrix <- cor(combined_cars[, c("Year_mfd","km_drv","Selling_price")])

ggcorplot(cor_matrix, method = "square", digits = 2, lab = T)
```
